Report

### 2014-12-07 MSLLReport.tex
2ed

\begin{document}
% coverpage
\begin{frame} % Cover slide
\titlepage
You can get this ppt from https://github.com/saiichi/LearnReport
\end{frame}
%section 1 Supervised Learning
\section{Supervised Learning}
%section 1.1 Discriminate Method
    \subsection{Review}
    %review slide
    \begin{frame}[options]
        \frametitle{Discriminate Method Review}

@@

                $$h_w(x) = g_w(w^Tx)=\left\{
                    \begin{array}{ll}
                        1  &\mbox{if $w^Tx \ge 0$}\\
                        -1  &\mbox{otherwise}
                    \end{array}
                    \right.
                $$
        \end{enumerate}
        %\end{block}

@@

			\end{itemize}
			\item Statistical decision theory: switch if $P(E=1|B=b) < \frac{1}{2}$
			\item $P(E=1|B=b)=\frac{P(B=b|E=1)P(E=1)}{P(B=b)}=\frac{1}{3}$
		\end{itemize}
	\end{frame}
	\begin{frame}
		\frametitle{Some concepts}
		\begin{itemize}
			\item The world is reduced to a set of random variables $x_1,\ldots,x_d$
			\begin{itemize}
				\item e.g. $(x_1,\ldots,x_{d-1}) $ is a feature vector, $x_d\equiv y$ is the class label.
			\end{itemize}
			\item Inference: given joint distribution $p(x_1,\ldots,x_d),$ compute $p(X_Q|X_E)$ where $X_Q\cup X_E \subseteq \{x_1,\ldots,x_d\}$
			\begin{itemize}
				\item e.g. $Q=\{x_d\},E=\{x_1,\ldots,x_{d-1}\},$ by the definition of conditional
				$$p(x_d|x_1,\ldots,x_{d-1})=\frac{p(x_1,\ldots,x_d)}{\sum_vp(x_1,\ldots,x_{d-1},x_d=v)}$$
			\end{itemize}
			\item Learning estimate $p(x_1,\ldots,x_d)$ from training data $\{x^{(1)},\ldots,x^{(m)}\}$
		\end{itemize}
	\end{frame}
	\begin{frame}
		\frametitle{What are graphical models?}
		\begin{itemize}
			\item Graphical model = joint distribution $p(x_1,\ldots,x_d)$
			\begin{itemize}
				\item Bayesian network
				\item Markov random field
			\end{itemize}
			\item Inference = $p(X_Q|X_E)$, in general $X_Q\cup X_E \subseteq \{x_1,\ldots,x_d\}$
			\item If $p(x_1,\ldots,x_d) not given, estimate it from data$
			\begin{itemize}
				\item parameter and structure learning
			\end{itemize}
		\end{itemize}
	\end{frame}
	\begin{frame}
		\frametitle{What are graphical models?(2)}
		\begin{itemize}
			\item Graphical model is the study of probabilistic models
			\item Just because there are nodes and edges doesn't mean it's a graphical model
			\item These are not graphical models:
			\begin{figure}

>>>>

\begin{document}
% coverpage
\begin{frame} % Cover slide
\titlepage
You can get this ppt from https://github.com/saiichi/LearnReport
\end{frame}
%section 1 Supervised Learning
\section{Introduction to Machine Learning}
\begin{frame}
	\frametitle{Machine Learning: The Original Motivation}
	Traditional Software Process:
	\begin{itemize}
		\item Interview the experts,
		\item Create an algorithm that automates their process.
	\end{itemize}
	Machine Learning process:
	\begin{itemize}
		\item Collect input-output examples from the experts.
		\item Learn a function to map from the input to the output.
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{Supervised Learning}
	\begin{itemize}
		\item Given: Training examples$(x_i,f(x_i))$ for some unknown function $f$.
		\item Find: A good approximation to $f$.
		\item Example Applications
		\begin{itemize}
			\item Handwriting recognition
			\begin{itemize}
				\item $x$ data from pen motion
				\item $f(x)$ letter of the alphabet
			\end{itemize}
			\item Disease Diagnosis
			\begin{itemize}
				\item $x$: properties of patient(symptoms, lab tests)
				\item $f(x)$: disease(or maybe, recommended therapy)
			\end{itemize}
			\item Face recognition
			\begin{itemize}
				\item $x$: bitmap picture of person's face
				\item $f(x)$: name of person
			\end{itemize}
			\item Spam detection
			\begin{itemize}
				\item $x$: email message
				\item $f(x)$: spam or not spam
			\end{itemize}
		\end{itemize}
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{Formal Setting}
	\begin{itemize}
		\item Training examples are drawn independently at random according to unknown probability distribution $P(x,y)$.
		\item The learning algorithm analyzes the examples and produces a classifier $f$.
		\item Given a new data point $(x,y)$ drawn from $P$, the classifier is given $x$ and predicts $\hat{y} = f(x)$.
		\item The loss $\mathcal{L}(\hat{y},y)$ is then measured.
		\item Goal of the learning algorithm: Find the $f$ that minimizes the expected loss
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{Formal Version of Spam Detection}
	\begin{itemize}
		\item $P(x,y)$: distribution of email message $x$ and their true labels $y$("spam" or "not spam").
		\item training sample: a set of email messages that have been labeled by the user.
		\item learning algorithm: logistic regression and so on.
		\item $f$: the classifier output by the learning algorithm.
		\item test point: A new email message $x$.
		\item loss function $\mathcal{L}(\hat{y},y)$
	\end{itemize}
\end{frame}

\begin{frame}
	\begin{itemize}
		\frametitle{Three main approaches to ML}
		\item Learn a classifier: a function $f$
		\item Learn a conditional distribution: a conditional distribution $P(y|x)$
		\item Learn the joint probability distribution: $P(x,y)$
		\item One example for each method
		\begin{itemize}
			\item A classifier: The Perceptron algorithm.
			\item A conditional distribution: Logistic regression.
			\item A joint distribution: Linear discriminant analysis.
		\end{itemize}
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{Linear Threshold Units}
	\begin{equation}
		h_w(x) = g_w(w^Tx)=\left\{
		\begin{array}{ll}
		1  &\mbox{if $w^Tx \ge w_0$}\\
		-1  &\mbox{otherwise}
		\end{array}
		\right.
	\end{equation}
	\begin{itemize}
		\item Where $x = (x_1,\ldots,x_n)$, $w = (w_1,\ldots,w_n)$.We assume that each feature $x_j$ and each weight $w_j$ is a real.
		\item For convenience, we transform $x = (x_1,\ldots,x_n)$ to $x=(-1,x_1,\ldots,x_n)$, then the parameter vector will then be $w=(w_0,w_1,\ldots,w_n)$
		\item Our goal is to find $w$
	\end{itemize}

\begin{frame}
	\frametitle{Geometrical View}
	\begin{itemize}
		\item Consider three training examples:\\
		$(<1.0,1.0>,+1),(<1.5,3.0>,+1),(<2.0,2.0>,-1)$
		\item We want a classifier that looks like the following:
		\begin{figure}
			\pgfimage[height=5cm ]{"image/35.png"}
		\end{figure}
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{ML == Optimization}
	\begin{itemize}
		\item Given:\\
			-A set of $N$ training examples\\
			\quad $\{x^{(1)},\ldots,x^{(N)}\}$
		\item Find:\\
			-The weight vector $w$ that minimizes the expected loss on the training data.
			\begin{equation}
				J(w) = \frac{1}{N}\sum_{i=1}^{N}\mathcal{L}(sgn(w^Tx^{(i)}),y_i)
			\end{equation}
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{Problem: Step-wise Constant Loss Function}
	\begin{figure}
		\pgfimage[height=5cm ]{"image/36.png"}
	\end{figure}
	Derivative is either $0$ or $\infty$
\end{frame}

\begin{frame}
	\frametitle{Approximating the expected loss by a smooth function}
	\begin{itemize}
		\item Simplify the optimization problem by replacing the original objective function by a surrogate loss function. For example, consider the hinge loss:
		\begin{equation}
			\widetilde{J}(w) = \frac{1}{N}\sum_{N}^{i=1}\max(0,1-y_iw^Tx^{(i)})
		\end{equation}
		\item When $y=1$
		\begin{figure}
			\pgfimage[height=3.5cm ]{"image/37.png"}
		\end{figure}
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{Gradient of the Hinge Loss}
	Let $\widetilde{J}_i(w) = \max(0,-y_iw^tx^{(i)})$
	$$\frac{\partial\widetilde{J}(w)}{\partial w_k} = \frac{\partial}{\partial w_k}\left(\frac{1}{N}\sum_{i=1}^{N}\widetilde{J}_i(w)\right)=\frac{1}{N}\sum_{i=1}^{N}\frac{\partial}{\partial w_k}\tilde{J}_i(w)$$
	\begin{eqnarray}
		\sum_{i=1}^{N}\frac{\partial}{\partial w_k}\tilde{J}_i(w) &=& \frac{\partial}{\partial w_k}\max\left(0,-y_i\sum_{j}w_jx_j^{(i)}\right) \\
		&=& \left\{
		\begin{array}{ll}
		0  &\mbox{if $w^Tx \ge w_0$}\\
		-yx_k^{(i)}  &\mbox{otherwise}
		\end{array}
		\right.
	\end{eqnarray}
\end{frame}

\begin{frame}
	\frametitle{Minimizing $\tilde{J}$ by Gradient Descent Search}
	\begin{figure}
		\pgfimage[height=3.5cm ]{"image/38.png"}
	\end{figure}
	\begin{itemize}
		\item Start with weight vector $w^{(0)}$
		\item Compute gradient $\nabla\tilde{J}(w^{(0)}) = \left(\frac{\partial\tilde{J}(w^{(0)})}{\partial w_0},\frac{\partial\tilde{J}(w^{(0)})}{\partial w_1},\ldots,\frac{\partial\tilde{J}(w^{(0)})}{\partial w_2}\right)$
		\item Compute $w^{(1)} = w^{(0)} - \eta \nabla\tilde{J}(w^{(0)})$
		\item Repeat until convergence
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{Logistic Regression}
	\begin{itemize}
		\item Learn the conditional probability $P(y|x)$
		\item Let $p(y|x;w)$ be our estimate of $P(y|x)$, where $w$ is a vector of adjustable parameters. Assume only two classes $y=0$ and $y=1$, and
		$$p(y=1|x;w) = \frac{\exp w^Tx}{1+\exp w^Tx}$$
		$$p(y=0|x;w) = 1-p(y=1|x;w)$$
		\item It is easy to show that this is equivalent to
		$$\log\frac{p(y=1|x;w)}{p(y=0|x;w)}=w^Tx$$
		\item In other words, the log odds of class 1 is a linear function of x
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{Why the exp function}
	One reason: linear function has a range from $-\infty$ to $+\infty$ and we need to force it to be positive and sum to 1 in order to be a probability:
	\begin{figure}
		\pgfimage[height=4cm ]{"image/39.png"}
	\end{figure}
\end{frame}

\begin{frame}
	\frametitle{Choosing the Loss Function}
	For probabilistic models, we use the log loss:
	$$\mathcal{L}(\hat{P}(y|x),y) =  \left\{
	\begin{array}{ll}
	-\log \hat{P}(y=1|x^{(i)})  &\mbox{if $y_i=1$}\\
	-\log \hat{P}(y+0|x^{(i)})  &\mbox{if $y_i=0$}
	\end{array}
	\right.$$ 
	\begin{figure}
		\pgfimage[height=5cm ]{"image/40.png"}
	\end{figure}
\end{frame}

\begin{frame}
	\frametitle{Maximum Likelihood Fitting}
	\begin{itemize}

	\end{itemize}
\end{frame}

\section{Supervised Learning}
%section 1.1 Discriminate Method
    \subsection{Review}
    %review slide
    \begin{frame}[options]
        \frametitle{Discriminate Method Review}

@@

                $$h_w(x) = g_w(w^Tx)=\left\{
                    \begin{array}{ll}
                        1  &\mbox{if $w^Tx \ge w_0$}\\
                        -1  &\mbox{otherwise}
                    \end{array}
                    \right.
                $$

        \end{enumerate}
        %\end{block}

@@

			\end{itemize}
			\item Statistical decision theory: switch if $P(E=1|B=b) < \frac{1}{2}$
			\item $P(E=1|B=b)=\frac{P(B=b|E=1)P(E=1)}{P(B=b)}=\frac{1}{3}$
		\end{itemize}
	\end{frame}
	\begin{frame}
		\frametitle{The probability explanation for ML}
		\begin{itemize}
			\item The world is reduced to a set of random variables $x_1,\ldots,x_d$
			\begin{itemize}
				\item e.g. $(x_1,\ldots,x_{d-1}) $ is a feature vector, $x_d\equiv y$ is the class label.
			\end{itemize}
			\item Inference: given joint distribution $p(x_1,\ldots,x_d),$ compute $p(X_Q|X_E)$ where $X_Q\cup X_E \subseteq \{x_1,\ldots,x_d\}$
			\begin{itemize}
				\item e.g. $Q=\{x_d\},E=\{x_1,\ldots,x_{d-1}\},$ by the definition of conditional
				$$p(x_d|x_1,\ldots,x_{d-1})=\frac{p(x_1,\ldots,x_d)}{\sum_vp(x_1,\ldots,x_{d-1},x_d=v)}$$
			\end{itemize}
			\item Learning estimate $p(x_1,\ldots,x_d)$ from training data $\{x^{(1)},\ldots,x^{(m)}\}$
		\end{itemize}
	\end{frame}
	\begin{frame}
		\frametitle{What are graphical models?}
		\begin{itemize}
			\item Graphical model = joint distribution $p(x_1,\ldots,x_d)$
			\begin{itemize}
				\item Bayesian network
				\item Markov random field
			\end{itemize}
			\item Inference = $p(X_Q|X_E)$, in general $X_Q\cup X_E \subseteq \{x_1,\ldots,x_d\}$
			\item If $p(x_1,\ldots,x_d) not given, estimate it from data$
			\begin{itemize}
				\item parameter and structure learning
			\end{itemize}
		\end{itemize}
	\end{frame}
	\begin{frame}
		\frametitle{What are graphical models?(2)}
		\begin{itemize}
			\item Graphical model is the study of probabilistic models
			\item Just because there are nodes and edges doesn't mean it's a graphical model
			\item These are not graphical models:
			\begin{figure}
